Non-Nested Two Phase Debiased Calibration Simulation

### 2024-04-29 final/reports/prelim_paper.tex
Added NN simulations. They are working well.

\begin{itemize}
  \item[1.] HT-estimator: $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{2i}},$
  \item[2.] Regression estimator (Reg): (FIXME)
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
    solves

@@

>>>>

\begin{itemize}
  \item[1.] HT-estimator: $\hat Y_{HT} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{2i}},$
  \item[2.] Regression estimator (Reg): Let $\hat Y_{NN, reg} = \hat Y_{HT} + (\hat
    X_c - \hat X_{2, HT}) \hat \beta_2$ where $\hat Y_{HT} = \sum_{i \in A_2} d_{2i}
    y_i$, $\hat X_c = W \hat X_{1, HT} + (1 - W)\hat X_{2, HT}$, $W = n_{1,eff}
    / (n_{1,eff} + n_{2, eff})$, $\hat X_{1, HT} = \sum_{i \in A_1} d_{1i} x_{i}$,
    $\hat X_{2, HT} = \sum_{i \in A_2} d_{2i} x_{i}$ and

    $$ \hat \beta_2 = \left(\sum_{i \in A_2} x_i x_i^T\right)^{-1} \sum_{i \in
    A_2} x_i y_i.$$

    Then $\hat{\bar Y}_{NN, reg} = \hat Y_{NN, reg} / N$.
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
    solves

@@

In addition to estimating the mean parameter $\bar Y_N$, we also construct
variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$.
% For each approach we give the variance estimate in Table~\ref{tab:varforms2}.

We run the simulation $1000$ times for each of these methods and compute the
Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
simulation replicate. The results are in Table~\ref{tab:nndc-mean}.

\begin{table}[ht!]
  \centering
\input{tables/nndcsim_mean.tex}
% This table was generated from ../../src/explore/proto-dctpmcmc.R
\caption{This table shows the results of Simulation Study 2. It displays the
Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
unbiasedness of each estimator for the estimators: HT, Reg, DC-Pop, and
DC-Est.}
\label{tab:nndc-mean}
\end{table}

### 2024-04-29 src/explore/20240425-nndcsim.qmd
Added NN simulations. They are working well.

@@

    ht <- ht_est(p1_df, p2_df, nrow(upop_df))
    reg <- reg_est(p1_df, p2_df, nrow(upop_df))

    return(
      tibble(Est = c("HT", "Reg"),
             Theta = c(ht[[1]], reg[[1]]),
             Var = c(ht[[2]], reg[[2]]),
             Iter = iter)
    )
  } %>% bind_rows()

@@

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest)

%>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest)

@@

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims))

%>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, MCVar, EstVar, VarVar, Ttest)

>>>>

#' This  function does the debiased calibration
#'
#' @param zmat - A n x p matrix
#' @param w1i - A vector of length n
#' @param d2i - A vector of length n
#' @param T1 - A vector of length p
#' @param qi - A scaler or a vector of length n
#' @param entropy - One of "EL", "ET"
solveGE <- function(zmat, w1i, d2i, T1, qi, entropy) {

  f <- function(lam, zmat, w1i, d2i, T1, qi, entropy, returnw = FALSE) {

    if (entropy == "EL") {
      w <- -d2i / drop(zmat %*% lam)
    } else if (entropy == "ET") {
      w <- d2i * exp(drop(zmat %*% lam))
    } else {
      stop("We only accept entropy of EL or ET.")
    }

    if (returnw) {
      return(w)
    } else {
      return(drop((w * w1i * qi) %*% zmat) - T1)
    }
  }

  j <- function(lam, zmat, w1i, d2i, T1, qi, entropy) {
    if (entropy == "EL") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) / drop(zmat %*% lam)^2) %*% zmat
    } else if (entropy == "ET") {
      res <- t(zmat) %*% diag((w1i * qi * d2i) * exp(drop(zmat %*% lam))) %*% zmat
    }

    return(res)
  }

  init <- c(rep(0, ncol(zmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, zmat = zmat, w1i = w1i, d2i = d2i,
          T1 = T1, qi = qi, entropy = entropy,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

  resw <- f(res$x, zmat, w1i, d2i, T1, qi, entropy, returnw = TRUE)

  if (!(res$termcd %in% c(1))) {
    return(NA)
  }

  return(resw)

dc_nn_ybar <- 
  function(p1_df,
           p2_df,
           pop_df,
           qi = 1,
           entropy = "EL",
           estT1 = TRUE,
           sampling = "srs-poisson") {

  d_vec <- 1 / (p2_df$pi2)
  if (entropy == "EL") {
    gdi <- -1 / d_vec
    gdi1 <- -1 / (1 / p1_df$pi2)
    gdip <- -1 / (1 / pop_df$pi2)
    gpinv <- d_vec^2
    gdi12 <- "-1 / d2i"
  } else if (entropy == "ET") {
    gdi <- log(d_vec)
    gdi1 <- log(1 / p1_df$pi2)
    gdip <- log(1 / pop_df$pi2)
    gpinv <- d_vec
    gdi12 <- "log(d2i)"
  } else {
    stop("This type of entropy has not yet been implemented.")
  }

  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
  zmat <- cbind(samp_X, gdi)
  w1i <- 1
  d2i <- 1 / p2_df$pi2

  if (estT1) {
    v_poisson <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * p2_df$Y^2) / nrow(pop_df)^2
    v_srs <- (1 / nrow(p2_df) - 1 / nrow(pop_df)) * var(p2_df$Y)
    d_eff <- v_poisson / v_srs
    n2_eff <- nrow(p2_df) / d_eff
    n1_eff <- nrow(p1_df)

    x1_hat <- 
      c(sum(p1_df$X1 * qi / p1_df$pi1),
        sum(p1_df$X2 * qi / p1_df$pi1))

    x2_hat <- 
      c(sum(p2_df$X1 * qi / p2_df$pi2),
        sum(p2_df$X2 * qi / p2_df$pi2))

    T1 <- c(nrow(pop_df),
            xc_hat[1],
            xc_hat[2],
            sum(gdip * qi))
  } else {
    w <- 1
    T1 <- c(nrow(pop_df),
            sum(pop_df$X1 * qi),
            sum(pop_df$X2 * qi),
            sum(gdip * qi))
  }

  dc_w <- solveGE(zmat, w1i, d2i, T1, qi, entropy)

  # Mean Estimation
  theta <- sum(p2_df$Y * dc_w * w1i) / nrow(pop_df)

  z_dc <- model.matrix(~1 + p2_df$X1 + p2_df$X2 + gdi)
  gam_dc <- 
    solve(t(z_dc) %*% diag(d2i * gpinv * qi) %*% z_dc,
          t(z_dc) %*% diag(d2i * gpinv) %*% p2_df$Y)

    if (estT1) {
      # Variance Estimation
      coefs <- as.numeric(gam_dc)
      u_mat <- 
      matrix(c(rep(1, nrow(p2_df)), p2_df$X1 * w, p2_df$X2 * w, gdi * qi), ncol = 4)
      eps <- (p2_df$Y - drop(u_mat %*% coefs))
      var_x <- 
      (1/nrow(p1_df) - 1/nrow(pop_df)) * var(matrix(c(p1_df$X1, p1_df$X2), ncol = 2))

      v1 <- 
      w^2 * matrix(coefs[2:3], nrow = 1) %*% var_x %*% matrix(coefs[2:3], ncol = 1)
      v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / nrow(pop_df)^2

      cov_xy <- c(sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[2] * 
                     (n2_eff) / (n1_eff + n2_eff) * p2_df$X1 * eps) / nrow(pop_df)^2,
                  sum((1 - p2_df$pi2) / p2_df$pi2^2 * coefs[3] * 
                     (n2_eff) / (n1_eff + n2_eff) * p2_df$X3 * eps) / nrow(pop_df)^2)
      v3 <- sum(cov_xy)

    } else {
      coefs <- as.numeric(gam_dc)
      u_mat <- 
      matrix(c(rep(1, nrow(p2_df)), p2_df$X1, p2_df$X2, gdi * qi), ncol = 4)
      eps <- (p2_df$Y - drop(u_mat %*% coefs))
      v2 <- sum((1 - p2_df$pi2) / p2_df$pi2^2 * eps^2) / nrow(pop_df)^2

      v_hat <- v2

  } else {
    v_hat <- NA
  }

@@

    ht <- ht_est(p1_df, p2_df, nrow(upop_df))
    reg <- reg_est(p1_df, p2_df, nrow(upop_df))
    dcpop <- dc_nn_ybar(p1_df, p2_df, upop_df, entropy = "EL", estT1 = FALSE)
    dcest <- dc_nn_ybar(p1_df, p2_df, upop_df, entropy = "EL", estT1 = TRUE)

    return(
      tibble(Est = c("HT", "Reg", "dcpop", "dcest"),
             Theta = c(ht[[1]], reg[[1]], dcpop[[1]], dcest[[1]]),
             Var = c(ht[[2]], reg[[2]], dcpop[[2]], dcest[[2]]),
             Iter = iter)
    )
  } %>% bind_rows()

@@

# Check for failures
filter(mc_res, is.na(Theta))

# Analyze Mean
mc_res %>%
  mutate(err = Theta - true_theta) %>%
  mutate(CI = abs(err) < qnorm(0.975) * sqrt(Var)) %>%
  group_by(Est) %>%
  summarize(Bias = mean(err, na.rm = TRUE),
            RMSE = sqrt(mean(err^2, na.rm = TRUE)),
            EmpCI = mean(CI, na.rm = TRUE),
            sdest = sd(Theta, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(Bias) / sqrt(sdest^2 / B_sims)) %>%
  select(-sdest) %>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  mutate(Est = case_when(
    Est == "dcest" ~ "DC-Est",
    Est == "dcpop" ~ "DC-Pop",
    Est == "HT" ~ "HT",
    Est == "Reg" ~ "Reg")
  ) %>%
  arrange(Ind) %>%
  select(Est, Bias, RMSE, EmpCI, Ttest)

@@

# Analyze Variance
mc_res %>%
  group_by(Est) %>%
  summarize(MCVar = var(Theta, na.rm = TRUE),
            EstVar = mean(Var, na.rm = TRUE),
            VarVar = var(Var, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Ttest = abs(MCVar - EstVar) / sqrt(VarVar / B_sims)) %>%
  mutate(Ind = case_when(
    Est == "dcest" ~ 4,
    Est == "dcpop" ~ 3,
    Est == "Reg" ~ 2,
    Est == "HT" ~ 1)
  ) %>%
  arrange(Ind) %>%
  select(Est, MCVar, EstVar, VarVar, Ttest)
